% Wayfinder manuscript — master document (LightsOut LaTeX approach, journal-adapted)
% Compile: latexmk -pdf main.tex   (or .\compile.ps1)
% Sections live in sections/*.tex (generated from the markdown by build_tex.py); edit prose there
% or in the markdown source and re-run build_tex.py.

\documentclass[11pt]{article}

% --- encoding / fonts (helvet sans, the LightsOut look) ---
\usepackage[utf8]{inputenc}
\DeclareUnicodeCharacter{03C9}{\ensuremath{\omega}}  % omega in a bib title
\DeclareUnicodeCharacter{2014}{---}                  % em-dash in a bib field
\usepackage[T1]{fontenc}
\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}\onehalfspacing
\setlength{\parindent}{1.6em}\setlength{\parskip}{3pt}
\tolerance=400\emergencystretch=3pt

% --- colours + section headers (LightsOut aesthetic; starred = unnumbered so the manual
%     "4.1b"-style numbers in titles match the prose's literal cross-references) ---
\usepackage[dvipsnames]{xcolor}
\definecolor{sectioncolor}{RGB}{0,0,0}      % black (standard manuscript headings)
\definecolor{subsectioncolor}{RGB}{0,0,0}
\definecolor{markercolor}{RGB}{0,0,0}
\usepackage{titlesec}
\titleformat{\section}{\normalfont\bfseries\large}{}{0em}{}
  [\vspace{-0.4em}\rule{\textwidth}{0.4pt}]
\titleformat{\subsection}{\normalfont\bfseries}{}{0em}{}
\titleformat{\subsubsection}[runin]{\normalfont\bfseries\itshape}{}{0em}{}[:]
\titlespacing*{\section}{0pt}{10pt}{4pt}
\titlespacing*{\subsection}{0pt}{6pt}{2pt}
\titlespacing*{\subsubsection}{0pt}{4pt}{0.5em}
\usepackage{indentfirst}  % indent the first paragraph after every heading too

% --- math / tables / figures / links ---
\usepackage{amsmath}
\usepackage{calc}  % defines \real{} for pandoc's longtable column-width arithmetic
\usepackage{graphicx,booktabs,longtable,array,caption}
\graphicspath{{figures/}}
\captionsetup{font=footnotesize,labelfont=bf,skip=3pt}
\usepackage[hidelinks]{hyperref}

% --- references (numbered superscript, LightsOut style; swap to author-year for Frontiers final) ---
\usepackage[numbers,super,sort&compress]{natbib}
\bibliographystyle{unsrtnat}

% --- pandoc helper macros (fragments use these) ---
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\providecommand{\passthrough}[1]{#1}
\providecommand{\pandocbounded}[1]{#1}
% pandoc longtables without captions + hyperref reference a counter literally named 'none'
\providecommand{\theHnone}{none.\arabic{none}}
\newcounter{none}

% --- title ---
\title{\vspace{-2em}\bfseries\color{sectioncolor}Receipt-backed prioritization for literature-based\\
  discovery using Perturb-seq evidence}
\author{Dayanjan Shanaka Wijesinghe}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Literature-based discovery (LBD) connects facts that are individually established but never joined, and
four decades of work have made it a prolific generator of hypotheses. Its persistent weakness is not
generation but triage: an LBD search returns far more candidate links than any laboratory can pursue, and
ranking by literature-rarity tracks obscurity as readily as importance.

\vspace{2pt}
We present \textbf{Wayfinder}, an approach that addresses this gap by pairing LBD generation with a
deterministic referee that evaluates each proposed \emph{gene $\rightarrow$ T-cell program $\rightarrow$
disease} hypothesis against a held experimental substrate --- a genome-scale CRISPRi Perturb-seq resource in
primary human CD4$^{+}$ T cells --- and returns, for each, a prioritization verdict with a receipt at every
hop. We keep two receipt classes distinct: the knockdown, effect, and program hops are backed by
experimental perturbation measurements, whereas the disease hop is a genetic-association nomination rather
than a claim of experimental causality. Two diagnostics define the referee's behaviour. It is willing to
answer a confident, receipt-backed \emph{no} --- a \emph{falsification} diagnostic --- and it distinguishes a
\emph{refuted} hypothesis from an \emph{untested} one, an explicit \emph{abstention}: a failed knockdown is
reported as untested, an artifact caught rather than a false negative recorded.

\vspace{2pt}
The entire loop --- generation, evidence integration, and provenance --- ran inside an agentic scientific
workbench in which an independent critic model enforced calibrated \emph{language} on the platform's own
output, and its analyses reproduced when the workbench was driven programmatically --- the pipeline's funnel
and ranking digit-for-digit, and the sensitivity panel byte-for-byte (replicable in principle, subject to
the platform's user interface). Applied to the CD4$^{+}$ resource, the approach culled 22{,}039
machine-generated hypotheses to a small set of receipt-backed survivors; a verdict ledger demonstrates that
the referee discriminates, and negative controls locate that discrimination precisely --- including the
finding that the disease hop's stringency is inherited from the substrate rather than supplied by the
referee. The worked example --- NAB2 $\rightarrow$ Th1/Th2 polarization $\rightarrow$ atopic eczema --- is a
literature-novel regulatory nomination whose sharpest artifactual confounder, a CRISPRi \emph{cis}-effect on
the adjacent \emph{STAT6}, we test against the study authors' own genome-wide data, where STAT6 is unmoved (no
detectable expression-level \emph{cis}-effect). The confounder we \emph{cannot} discharge ---
whether the eczema label is linkage-inherited within the 12q13 locus --- we foreground rather than claim
resolved; an independent cross-model replication reproduced the finding. The contribution is the method and what it found:
\textbf{receipt-backed prioritization of machine-generated hypotheses, with explicit abstention and
falsification diagnostics} --- not a demonstration of predictive correctness.

\vspace{4pt}\noindent\textbf{Keywords:} literature-based discovery; hypothesis prioritization; Perturb-seq;
CRISPRi screens; T-cell polarization; agentic language models; falsification; abstention.
\end{abstract}

\section*{1. Introduction}\input{sections/01_introduction}
\section*{2. Background}\input{sections/02_background}
\section*{3. Materials and Methods}\input{sections/03_methods}
\section*{4. Results}\input{sections/04_results}
\section*{5. Discussion}\input{sections/05_discussion}

\section*{Acknowledgements}
This work began at the \textbf{\emph{Built with Claude: Life Sciences}} hackathon organized by
\textbf{Cerebral Valley}, whose team we thank for convening the event. The entire study is a re-analysis of a
public resource: we are grateful to the authors of the genome-scale CD4$^{+}$ T-cell Perturb-seq dataset ---
Zhu, Dann, and colleagues in the \textbf{Marson laboratory} --- for depositing their data and aggregated
supplementary tables openly \cite{zhu2025}, without which none of this would have been possible. The
generation, evidence integration, and self-audit were carried out inside \textbf{Claude Science}, Anthropic's
agentic scientific workbench, and the reasoning and language models throughout are Anthropic's Claude family
(Section 3.1). We gratefully acknowledge \textbf{Anthropic} for a Claude Max account and the API credits that
made the agentic experiments and their cross-model verification feasible. Responsibility for all claims,
and for the calibrated language the paper argues for, rests with the author.

\section*{Data and code availability}
The Perturb-seq resource re-analysed here is public \cite{zhu2025}. All prioritization runs against its
aggregated supplementary tables; the one genome-wide confounder check reads the study's deposited
differential-expression matrix by byte-range from public object storage. The Wayfinder implementation,
analysis scripts, and figure-generation code are released open-source (MIT).

\bibliography{../../../references/references}

\end{document}
