We report six results. The funnel and verdict ledger show the referee discriminating across a spread of
worked examples (\S{}4.1), and two negative controls plus a rank-stability check probe how much of that
discrimination is a property of the data versus an artifact of the setup --- with one control returning a
deliberately honest, non-obvious answer (\S{}4.1b) --- and a gate-threshold grid shows the flagship's rank is
stable wherever it stays eligible (\S{}4.1c). The confident \emph{no} and the quality-control
catch are examined on their own (\S{}4.2), because they are the point. NAB2 is then followed hop by hop as the
worked hero (\S{}4.3), and its sharpest artifactual confounder is falsified against the study authors' own
genome-wide data (\S{}4.4), while the one confounder the substrate \emph{cannot} discharge --- the 12q13
GWAS-label provenance --- is foregrounded rather than buried (\S{}4.4b). Finally we show the entire loop reproduced natively inside the agentic workbench,
audited by an independent critic model (\S{}4.5), and replicated by a second vendor's models (\S{}4.6).

\subsection*{4.1 The funnel and the verdict ledger}\label{the-funnel-and-the-verdict-ledger}

At the 8-hour stimulation condition, the generator built a disease-answer-free universe of \textbf{3,935}
knockdown-gated, program-significant genes, from which the gate admitted \textbf{22,039} eligible
(gene, disease) pairs. Running the deterministic referee over those pairs returned \textbf{43} with a positive
disease hop, of which \textbf{30} held the full chain cleanly (nonzero positive effect); the remainder split into
10 effect-zero \emph{supported-weak}, 3 off-target \emph{supported-flagged}, and 1 \emph{refuted-effect}, with the other
\textbf{21,995} pairs \emph{refuted for the specific disease} (Table 3). The generate-to-cull ratio is the headline
shape: 22,039 machine-generated questions reduced by data receipts to 30 clean survivors, none asserting
more than a table value supports.

We are deliberately careful about what this funnel does and does not demonstrate, and we restate the two
honesty caveats here rather than defer them to Limitations. First, the universe \textbf{A} is preselected on
program-significance, so \emph{within the funnel} the program hop cannot fail (\texttt{refuted\_program\ $\equiv$\ 0} by
construction); the dominant cull among eligible pairs is therefore disease-specificity, not broad four-hop
falsification. The program hop discriminates in an individual triple's receipt (a reported quantity --- NAB2's
Ota \emph{z} = 7.71), not as an independent filter on funnel survivors. Second, the ``43'' is a \textbf{joint} product
of the literature-novelty gate and the referee, not a referee-only count: the referee alone supports
\textbf{395 of 47,220} pairs, and the novelty gate culls 395 $\rightarrow$ 43. The honest reading is ``referee-supported
\emph{among literature-eligible} pairs.'' Accordingly we present the ledger below as a \textbf{demonstration that the
referee discriminates} --- that on worked examples it will support, decline to test, and refute, each with a
receipt --- not as a measured precision or recall.

The verdict ledger (Table 2) places five genes across the full verdict spectrum, each with its per-hop
receipt:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gene $\rightarrow$ disease (condition)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Verdict
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Deciding receipt
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{NAB2} $\rightarrow$ atopic eczema (Stim8hr) & \textbf{supported} & full chain holds; eczema clusters OR 3.90 / FDR 0.0028 and OR 3.43 / FDR 0.0224 (\S{}4.3) \\
\textbf{EGR2} $\rightarrow$ asthma (Stim8hr) & \textbf{supported} & 2/2 guides signif; effect -11.06, 854 downstream DE; member of 34 disease clusters, asthma OR 20.4 / FDR 5$\times$10\textsuperscript{-}\textsuperscript{5} \\
\textbf{NUDT1} $\rightarrow$ type 1 diabetes (Stim8hr) & \textbf{supported-weak} & full chain holds but a trivial effect (4 downstream DE) --- the \emph{only} pure-disjoint (zero-literature) survivor \\
\textbf{IL2} $\rightarrow$ (Rest) & \textbf{untested} & knockdown-QC fails: 0/2 guides significant; target barely expressed at Rest, nothing to knock down \\
\textbf{SLC1A5} $\rightarrow$ type 1 diabetes (Stim8hr) & \textbf{refuted} & in 9 disease cluster gene-sets, but none reach FDR \textless{} 0.05 (best FDR 0.054, OR 2.73) \\
\end{longtable}
}

Two features of the ledger carry the thesis. The \emph{untested} verdict (IL2) is not a missing entry or a
failure to compute --- it is the referee refusing to read an uninterpretable knockdown as a biological
result. And the \emph{refuted} verdict (SLC1A5) is a plausible metabolite-transporter $\rightarrow$ autoimmunity link that
the enrichment data decline to support: a confident, receipt-backed \emph{no}. NUDT1 illustrates the opposite
edge --- a hypothesis that clears every hop yet is foregrounded only weakly, because its effect is trivial;
this is exactly why the pipeline \emph{ranks} by a novelty-plus-effect score rather than hard-gating on
zero-literature absence, which would reward obscurity (\S{}3.2). We note that these examples are drawn to span
the verdict space, not sampled to estimate a rate; \S{}4.1b supplies a disease-hop support-\emph{rate} control over
the full A$\times$C space --- not a ledger accuracy rate, nor a claim of precision or recall.

\subsection*{4.1b Sensitivity: two negative controls and rank stability}\label{b-sensitivity-two-negative-controls-and-rank-stability}

To move ``the referee discriminates'' from assertion to a small measured result, we ran three checks as a
bounded kernel analysis inside the workbench (\S{}3.4).

\textbf{Control 1 --- the QC gate does its job.} We applied the referee to every gene whose knockdown \emph{fails} QC
at the 8-hour condition (no guide reaching a significant knockdown call in T4): \textbf{2,430} genes. All 2,430
returned \emph{untested} at hop 0 --- a \textbf{100\%} rate, with no leakage into a spurious \emph{refuted} or \emph{supported}
verdict. The gate converts failed experiments into an honest ``cannot say,'' exactly as designed, and does so
without exception across the full failed-knockdown set.

\textbf{Control 2 --- the disease hop's stringency is substrate-inherited, and its pass count is label-dependent.}
Across the referee's native pair space of all \textbf{47,220} (gene $\in$ A) $\times$ (disease $\in$ C) combinations at the
8-hour condition, the disease hop supports only \textbf{406 (0.86\%)} and refutes the remaining \textbf{99.14\%}. That
stringency, however, is not the referee's own doing: under \textbf{2,000} permutations of the disease labels
across the enrichment rows, the null still passes only \textbf{467.7 $\pm$ 10.9} pairs --- a null pass rate of
\textbf{0.99\%} --- so even randomly-labelled pairs clear FDR \textless{} 0.05 barely more than one percent of the time. The
near-total refutation is therefore \textbf{inherited from the enrichment study's FDR family} --- a
substrate-provenance property we flag in Limitations (\S{}5.3), and one an adversarial replication
independently noted --- not a discrimination we should credit to the referee. What the permutation \emph{does}
establish is that the pass count is \textbf{label-dependent}: the observed 406 sits \textbf{5.6 standard deviations
below} the null mean (lower-tail permutation \emph{p} $\approx$ 5$\times$10\textsuperscript{-}\textsuperscript{4}; the \emph{upper}-tail \emph{p} is 1.0 --- that is, the hop
does \textbf{not} pass \emph{more} often than chance, so we explicitly do not claim a ``rarer-than-chance''
selectivity). The direction --- observed \emph{below} null --- is a feature of this restricted twelve-disease
permutation setup; the one component we actually measured, a per-gene concentration effect, is small (a
passing gene carries \textbf{2.39} distinct disease labels under the true assignment versus \textbf{2.44} under
shuffling), so we report the figures but do not interpret the mechanism of the direction further. The load-bearing reading is the
one we can defend on both tails: the disease hop is a \textbf{stringent, label-dependent nomination filter whose
stringency is substrate-inherited} --- not a demonstration of the referee's own discriminating power. (This
control speaks to the disease hop; it does not re-open the program-hop tautology of \S{}3.3, which is a
construction property, not a chance one.)

\textbf{Rank stability under alternative objective weights.} The ranking objective (\S{}3.2) has three free weights
($\beta$ on effect, \emph{w} on direct co-mention, \emph{w\textsubscript{2}} on curated association; default $\beta$ = 1, \emph{w} = 1, \emph{w\textsubscript{2}} = 3). The
weights set \emph{priority} among eligible survivors, not the supported/refuted verdict, which is
weight-independent by construction --- but a headline example that is rank-4 only at one weight setting would
be a fragile choice of hero. We recomputed the ranking over a \textbf{27-point grid} crossing $\beta$ $\in$ \{0.5, 1, 2\},
\emph{w} $\in$ \{0.5, 1, 2\}, and \emph{w\textsubscript{2}} $\in$ \{1, 3, 5\}, recovering each survivor's weight-independent score component from
its default-weight value and re-ranking under every setting (the default setting reproduces NAB2's rank of
4, confirming the recovery). NAB2's rank ranged from \textbf{1 to 8} across the grid, with a \textbf{median of 4}, and
it stayed within the top five survivors under \textbf{24 of the 27} settings (89\%). No survivor's \emph{verdict}
changed, since verdicts do not depend on the weights. The worked hero is robust to the one human judgment
(\S{}3.2) in an otherwise mechanical pipeline.

\subsection*{4.1c Threshold robustness: survivors and the flagship rank across the gate}\label{threshold-robustness-survivors-and-the-flagship-rank-across-the-gate}

Section 4.1b varied the ranking \emph{weights}; the eligibility \emph{gate} (\S{}3.2) has its own three
thresholds --- the literature-percentile floor \texttt{ab\_gate\_pct} (default the median, 0.50), the
program--disease co-mention floor \texttt{min\_bc} (default 3), and the novelty ceiling \texttt{tau} on
curated association (default 0.10). We swept a \textbf{27-point grid} crossing \texttt{ab\_gate\_pct} $\in$
\{0.25, 0.50, 0.75\}, \texttt{min\_bc} $\in$ \{2, 3, 5\}, and \texttt{tau} $\in$ \{0.05, 0.10, 0.20\},
recomputing the funnel offline for every cell. Because a pair's referee verdict does not read the gate, the
verdicts are computed once over the full 47,220-pair space and re-filtered per cell (the default cell
reproduces the 30 clean survivors and NAB2's rank of 4 exactly). We read three things from the grid, as a
diagnostic of the machinery's response to its thresholds, not as a validation.

\emph{The survivor count behaves as a gate should.} The clean full-chain survivor count ranges from
\textbf{7} at the strictest cell to \textbf{74} at the loosest, rising as the gate is relaxed, with the
default setting (30) in mid-range --- the gate admits more candidates as it loosens, without discontinuity.

\emph{The flagship's rank is stable wherever it is eligible.} Whenever NAB2 $\rightarrow$ atopic eczema clears
the gate, it ranks in a tight band --- \textbf{1, 4, or 5} across every such cell, median \textbf{4}, never
worse than fifth --- echoing the rank-4 median of the weight sweep (\S{}4.1b). Its \emph{verdict} is
invariant: the referee supports the pair in all 27 cells.

\emph{But eligibility depends on the literature floor, and we say so plainly.} NAB2 survives as a clean
nomination in \textbf{18 of 27} cells, and the nine exceptions are exactly the nine at
\texttt{ab\_gate\_pct}~=~0.75 --- independent of \texttt{min\_bc} and \texttt{tau}. NAB2's A--B program-literature
co-mention sits between the median and the 75th percentile of the universe, so a floor at the 75th percentile
removes it from the \emph{candidate pool} --- not because its receipts fail (they do not) but because a
near-novel gene, by construction, carries only moderate program-literature. This is the honest cost of
ranking near-novelty: the flagship is robust to the two disease-side thresholds (\texttt{min\_bc},
\texttt{tau}) but sensitive to how aggressively the literature floor prunes understudied genes, a dependence
a reader should weigh alongside the nomination. The clean-survivor \emph{sets} overlap the default by a
Jaccard of \textbf{0.41} at the median cell (minimum 0.23) --- a gate that reshapes the candidate pool around
a stable core. (This grid is deterministic and offline; recomputing the survivor sets and NAB2's rank reads
only cached literature counts.)

\subsection*{\texorpdfstring{4.2 The confident \emph{no}, and the quality-control catch}{4.2 The confident no, and the quality-control catch}}\label{the-confident-no-and-the-quality-control-catch}

The moat of this approach is a referee that will answer \emph{no} --- and that distinguishes a \emph{refuted}
hypothesis from an \emph{untested} one. Both behaviors appear in the ledger, and both are worth stating plainly.

IL2 is the quality-control catch. IL2 is a canonical, biologically central T-cell gene that appears in both
the Th program and the disease gene-sets; a naive pipeline that read the empty downstream result of a
failed knockdown would call it ``no effect'' and record a false negative. The referee instead halts at hop 0:
0 of 2 guides reach a significant knockdown (guide mean expression 0.031 vs.~non-targeting-control 0.036 ---
the target is barely expressed at Rest, so there is nothing to knock down), and the verdict is \emph{untested ---
knockdown failed QC}. An artifact is caught, not mistaken for biology.

SLC1A5 is the correctly-sourced refutation. A metabolite transporter with a plausible metabolic-gene $\rightarrow$
autoimmunity story, SLC1A5 clears the QC gate (1/2 guides significant) and shifts the Th1/Th2 program, yet
its disease hop \emph{refutes}: it appears in nine autoimmune-disease cluster gene-sets but none reaches
FDR \textless{} 0.05 (best FDR 0.054, type 1 diabetes, OR 2.73). The enrichment data decline the link. This is the
verdict a confirmation-shaped pipeline never produces, and it is deterministic --- the \emph{no} is computed, not
argued.

\subsection*{4.2b Out-of-funnel discrimination: the referee's own edge, quantified}\label{out-of-funnel-discrimination-the-referees-own-edge-quantified}

Section 4.1b established that the disease hop's near-total refutation rate is \emph{substrate-inherited} --- a
property of the enrichment study's FDR family, not the referee's own doing. That invites the fair question a
referee report pressed: once the substrate-inherited disease hop is set aside, how much does the referee
\emph{itself} discriminate? Its own hops are the knockdown-QC gate (hop 0) and the effect and program checks
(hops 1--2), each a deterministic computation the referee owns. We quantify their discrimination two ways,
both offline and outcome-independent.

\emph{On arbitrary genes.} We applied the referee to \textbf{all 11,415} genes perturbed at the 8-hour
condition --- the full out-of-funnel population, not the pre-gated universe --- pairing each with an arbitrary
disease (the referee's hops 0--2 are disease-independent). The referee's own hops cull \textbf{16.9\%} of
them: \textbf{16.8\%} (1{,}914 genes) return \emph{untested} because the knockdown fails QC --- genes the
substrate cannot speak to --- with only a small remainder refuted at the effect or program hop. The
knockdown-QC gate is thus the referee's dominant own-edge discriminator (one arbitrary gene in six cannot be
tested here), while the effect gate, lenient by design (any nonzero downstream signal passes), rarely fires
on unselected genes. This is the honest shape of the referee's own \emph{no}: it is real, and it is
concentrated in the quality-control catch rather than spread across the chain.

\emph{On a frozen nominator's best guesses (hard negatives).} A stronger test must not choose genes by the
outcome being evaluated. We froze a naive curated-association nominator --- for each of the 12 diseases, the
top 50 perturbed genes by Open Targets association score, selected with no reference to the referee --- and
applied the referee to the resulting \textbf{600} association-plausible (gene, disease) nominations. The
verdicts decompose exactly along the line the reviews asked us to draw: \textbf{15.7\%} are culled at the
referee's \emph{own} hops (\textbf{75} \emph{untested}, 8 \emph{refuted-effect}, 11 \emph{refuted-program}),
\textbf{67.3\%} are \emph{refuted for the specific disease} (the substrate-inherited hop, \S{}4.1b), and only
\textbf{17.0\%} are supported. The own-hop culls are the hard negatives --- genes a database-driven nominator
would confidently advance that the data cannot support here --- and they include strongly-associated flagship
pairs: \emph{IL36RN} $\times$ psoriasis (association 0.82), \emph{TREX1} $\times$ systemic lupus (0.78), and
\emph{PADI4} $\times$ rheumatoid arthritis (0.68), each returned \emph{untested} because its knockdown failed
QC in this screen. A curated-association ranking, however strong, cannot tell you whether the perturbation is
even measurable; the referee's own hops can, and they flag roughly one in six of the top nominations. This is
the discrimination the referee supplies on its own --- reported separately from, and not conflated with, the
disease hop's substrate-inherited stringency.

\subsection*{4.3 NAB2 $\rightarrow$ Th1/Th2 $\rightarrow$ atopic eczema, hop by hop}\label{nab2-th1th2-atopic-eczema-hop-by-hop}

The highest-ranked near-novel survivor is \textbf{NAB2}, near-novel by an operational criterion --- low direct
literature co-mention (\texttt{ac\_lit} = 6) and no curated Open Targets association (\texttt{ac\_known} = 0.038) --- rather
than strict Swanson-style A--C absence. The data support NAB2 as a Th1/Th2 regulator with a receipt at every
hop (Figure 4):

\begin{itemize}
\tightlist
\item
  \textbf{Hop 0 --- knockdown QC.} 2/2 guides significant, best adjusted \emph{p} $\approx$ 1$\times$10\textsuperscript{-}\textsuperscript{1}\textsuperscript{6} (guide expression 0.056
  vs.~non-targeting control 0.567). The perturbation worked.
\item
  \textbf{Hop 1 --- effect.} On-target knockdown, effect -16.9 (-16.88 exact), \textbf{301 downstream differentially expressed genes},
  no off-target flag. A large, clean transcriptional consequence.
\item
  \textbf{Hop 2 --- program.} Th2-associated in the Ota contrast \cite{ota2021} (\emph{z} = 7.71, adjusted \emph{p} = 1.96$\times$10\textsuperscript{-}\textsuperscript{1}\textsuperscript{3},
  log-fold-change +0.63; polarity marker-checked); not significant in the H\"{o}llbacher contrast \cite{hollbacher2020} --- one of
  two contrasts, reported rather than hidden.
\item
  \textbf{Hop 3 --- disease.} Member of two atopic-eczema-enriched modules: OR 3.90 (FDR 0.0028) and OR 3.43
  (FDR 0.0224). These are genome-wide functional immune modules (significant clusters 90 and 100 ---
  BACH2, BCL6, IRF4, CD28, IL4, IL10), not a 12q13 locus artifact.
\end{itemize}

The precisely scoped claim is: \emph{consistent with a re-derived NAB2 $\rightarrow$ Th1/Th2 $\rightarrow$ atopic-eczema chain the literature
has not made.} We keep the two receipt classes explicit even for the hero --- the knockdown, effect, and
program hops rest on experimental perturbation measurements; the atopic-eczema link is a \textbf{genetic-
association nomination} (an Open Targets GWAS-based disease label, without colocalization or LD control),
not an expression claim or a proof of disease causality. An independent four-agent literature audit found
zero papers connecting NAB2 either to Th1/Th2 polarization or to atopic eczema --- the finding is novel in
the literature --- and the same audit surfaced the confounder we take up next.

\subsection*{\texorpdfstring{4.4 Falsifying the hardest confounder: the STAT6 \emph{cis}-effect}{4.4 Falsifying the hardest confounder: the STAT6 cis-effect}}\label{falsifying-the-hardest-confounder-the-stat6-cis-effect}

NAB2 sits \textasciitilde1.9 kb from \emph{STAT6}, the master type-2/atopic regulator, so the sharpest artifactual explanation
of NAB2's \emph{perturbation} signal is a CRISPRi \emph{cis}-effect: a guide targeting NAB2 could inadvertently
repress the adjacent STAT6, and the observed Th2/eczema signal could be STAT6's, not NAB2's. We test this
directly against the study authors' own deposited genome-wide differential-expression matrix (the 16.8 GB
\texttt{GWCD4i.DE\_stats.h5ad}, read lazily from public object storage by byte-range, no download). Under NAB2
knockdown at Stim8hr:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2000}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2667}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2667}}
  >{\centering\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2667}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gene (under NAB2-KD)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
log\textsubscript{2}FC
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
adj. \emph{p}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
moved?
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{STAT6} (neighbor under test) & \textbf{+0.087} & \textbf{0.788} & no --- \textbf{unmoved} \\
\textbf{NAB2} (self / on-target) & \textbf{-3.078} & 7.2$\times$10\textsuperscript{-}\textsuperscript{6}\textsuperscript{0} & yes --- knockdown worked \\
\end{longtable}
}

Of 10,282 measured genes, 302 are significantly moved by NAB2 knockdown ($\approx$ the referee's effect count of
301); STAT6 is not among them, ranking 5,444/10,282 by absolute log-fold-change --- the less-affected half. A
\emph{cis}-artifact would push STAT6 \emph{down}; it does not move. Three further checks corroborate this --- drawn from
the study's own QC fields and our finding analysis rather than the genome-wide DE row above: NAB2 and STAT6
share zero perturbation-effect clusters (their knockdowns are not equivalent); NAB2 clears the study's own
reproducibility bar (cross-guide/donor R 0.74); and the authors' own off-target flag --- a transcription
start site within 10 kb showing significant down-regulation --- is False for NAB2, meaning their pipeline
already found no \emph{cis} down-regulation of any near neighbor, STAT6 included. We therefore report the
\emph{cis}-artifact as \textbf{refuted}: the signal is consistent with NAB2 perturbation rather than STAT6 bleed. We
scope the claim precisely --- this strengthens the case that the \emph{perturbation} signal is NAB2-specific; it
does not prove the disease link, which remains a genetic-association nomination.

\subsection*{\texorpdfstring{4.4b The confounder we cannot discharge: the 12q13 atopy locus}{4.4b The confounder we cannot discharge: the 12q13 atopy locus}}\label{the-confounder-we-cannot-discharge-the-12q13-atopy-locus}

The STAT6 falsification addresses one of \emph{three} distinct ways the 12q13 locus could manufacture NAB2's
eczema signal, and we separate them deliberately because only two are answerable with our data --- and we do
not want the one we can falsify to stand in for the one we cannot. NAB2 lies within the 12q13
atopic-dermatitis susceptibility locus, \textasciitilde1.9 kb from \emph{STAT6}. \textbf{(i) A CRISPRi \emph{cis}-effect}
--- the NAB2 guide silently repressing the adjacent STAT6, so the transcriptional signal is STAT6's --- is
\textbf{falsified}: STAT6 is unmoved under NAB2 knockdown (\S{}4.4). \textbf{(ii) A cluster-membership
artifact} --- NAB2's eczema modules being a 12q13-colocated gene set rather than genome-wide functional
modules --- is \textbf{rejected}: the significant modules (clusters 90 and 100) are genome-wide functional
immune modules (BACH2, BCL6, IRF4, CD28, IL4, IL10), with STAT6 absent (\S{}4.6). \textbf{(iii) An
LD-inherited disease label} we \textbf{cannot} settle. The disease label on those modules is a GWAS-based
genetic association, assigned without colocalization or linkage-disequilibrium analysis; because NAB2 sits
inside the 12q13 atopy locus, its eczema association could in principle be linkage-inherited from the locus's
lead signal --- plausibly STAT6's --- rather than being NAB2-specific. Our substrate carries the aggregated
enrichment result, not the variant-level GWAS, so it \emph{cannot} perform the colocalization that would
separate a genuinely NAB2-specific association from an LD shadow. We therefore \textbf{foreground this as the
flagship nomination's key open question, and explicitly do not claim it discharged.} The perturbation
evidence for NAB2 as a Th1/Th2 regulator is receipt-backed and confounder-tested; the further step from that
regulatory role to \emph{atopic eczema specifically} rests on a genetic-association label whose locus
provenance remains unresolved. Resolving it requires variant-level colocalization against the 12q13
atopic-dermatitis GWAS --- outside this substrate, and precisely the disease-hop analysis the loop nominates
next (\S{}5.2, \S{}5.3). Stating the boundary this plainly is itself the calibrated-language discipline the
paper argues for: the honest verdict here is \emph{a receipt-backed regulatory nomination with an unresolved
disease-label provenance}, not a settled NAB2$\rightarrow$eczema link.

\subsection*{4.5 Native reproduction and self-audit in Claude Science}\label{native-reproduction-and-self-audit-in-claude-science}

The entire loop --- generation, evidence integration, and the assembly of its own provenance --- was run inside the
agentic workbench, at a total run cost of \textasciitilde\$6.41, with the model identities, per-step costs, and reviewer
checks drawn from the workbench's own audit store. We keep the three liveness claims distinct (\S{}3.4). \emph{Full-
scale reproduction}: the real generator ran unchanged over the entire 3,935-gene universe inside the
workbench and reproduced the exact funnel (3,935 $\rightarrow$ 22,039 $\rightarrow$ 43 $\rightarrow$ 30) and NAB2's rank of 4, but against a
workbench-local cache of previously captured literature/association responses under a guard that raises on
any live network call (cache delta zero) --- a faithful recomputation, not a live crawl. \emph{Live authorship}: a
separate 12-gene from-scratch run, given only the method, had the workbench write its own generator and make
live Europe PMC and Open Targets calls; NAB2's effect was too low to rank in that small run's top twelve, so
it did not appear --- nothing was cherry-picked --- and independent re-querying of the workbench's recorded
queries against the live APIs returned identical counts. \emph{Cross-family independence} is kept external by
design (\S{}4.6).

The self-audit is the part that mirrors the paper's own thesis (Figure 5). The author model (an Opus-class
model) assembled the end-to-end receipt chain; an independent reviewer model (a Sonnet-class model),
invoked at separate checkpoints, verified every reported number against the underlying artifacts and
enforced calibrated language on the manuscript-facing output. It produced four verification checks: two
confirming that every cited receipt and the final ``what ran where'' table matched the source of truth, and ---
the telling one --- a check that \textbf{flagged the words ``validated'' (in a title) and ``definitive'' (in a heading)
as calibrated-language violations, both of which were then removed}. The platform corrected its own
overstatement. We characterize this independence honestly: it is \emph{role, model, and checkpoint} independence
--- a distinct reviewer model at distinct verification points within a single model family --- not independence
across vendors. It is also why manuscript-facing verdicts cite the post-review receipt rather than the raw
generation log, which still carries a pre-critic ``validated'' string in a free-text field: that pre- to
post-critic difference \emph{is} the self-audit.

\subsection*{4.6 Independent cross-model replication}\label{independent-cross-model-replication}

Cross-family independence was supplied from outside the workbench, by a five-member replication lab: three
Opus-class agents and two Codex (a different vendor's) agents, each re-deriving the finding from the raw
supplementary tables under an adversarial mandate, with two members performing clean-room re-
implementations that imported none of our pipeline code. The verdict was a \textbf{unanimous pass}. Across the
lab, every headline number was independently reproduced to the unit --- not every member re-checked every
number, but each number was re-derived from the raw CSVs by at least one clean-room member: the NAB2 receipt (2/2 guides,
effect -16.88, 301 downstream DE, Ota \emph{z} +7.71, eczema clusters OR 3.90/3.43), the full funnel
(3,935 $\rightarrow$ 22,039 $\rightarrow$ 43 $\rightarrow$ 30, with the 30/10/3/1 class split and the single pure-disjoint survivor), and the
answer-free construction of the universe.

The replication is worth reporting precisely \emph{because} it was not silent agreement: the adversarial pass
caught and corrected real errors. It found a cluster-ID misalignment in our confounder script --- sorted
cluster IDs printed against row-order FDRs had mislabeled NAB2's significant eczema modules as 74/90 when
the true significant modules are 90 and 100 (cluster 74 is non-significant, FDR 0.52), and the locus test
had been run on the wrong cluster; the corrected run confirms modules 90 and 100 are genome-wide functional
immune modules with STAT6 absent, which rejects the \emph{cluster-membership} locus artifact --- though not the
distinct question of whether the GWAS disease \emph{label} is LD-inherited from the 12q13 atopy locus, which the
substrate cannot settle and which we leave open (\S{}5.3). It corrected an effect-size overstatement (``\textasciitilde8$\times$ stronger'' is the
log-fold-change ratio; the \emph{z}-ratio is \textasciitilde3$\times$). And it reframed the distinctness argument onto its strongest
support --- cluster co-membership, magnitude, and guide-specificity --- rather than the shared disease profile,
which if anything \emph{aids} the confounder. That a hostile independent re-derivation reproduced every number
and improved the framing is the strongest available evidence that the finding is a property of the data,
not of our implementation.
