\subsection*{5.1 What the agentic loop changes}\label{what-the-agentic-loop-changes}

We opened with a critique the field has leveled at itself --- that literature-based discovery ``generates an
enormous number of hypotheses, almost none of which ever get followed up.'' Wayfinder is a response to that
critique, and it is worth being exact about \emph{which} part of it we answer. We do not claim to have solved
experimental follow-up; the wet-lab confirmation of any surviving hypothesis remains future work, as it
must. What the approach changes is the step immediately before that: \textbf{triage}. Of 22,039
machine-generated gene$\rightarrow$program$\rightarrow$disease hypotheses, the deterministic referee reduced the set to a few
receipt-backed survivors --- and, crucially, attached to each survivor a per-hop receipt and to each
rejection a reason. ``Which of these thousands is worth a bench's time?'' stops being a question answered by
literature-rarity heuristics (which, as we showed, track obscurity as readily as importance) and becomes a
question answered by a data test with a verdict a reader can inspect. That is a narrower claim than
``closing the LBD loop,'' and we prefer the narrower claim because it is the one the evidence supports.

Two properties of the loop do the work. The first is the willingness to answer \emph{no}, and to distinguish a
\emph{refuted} hypothesis from an \emph{untested} one. A generator with no discriminating filter turns one flood into
another; a filter that only ever confirms is not a filter. The knockdown-QC gate is the sharpest expression
of this --- a failed perturbation is reported as \emph{untested}, never as a false ``no effect'' --- and it is why the
approach can be trusted to reject as well as to nominate. The second is that the entire loop, generation
through self-audit, ran inside an agentic workbench in which an independent reviewer model enforced
calibrated language on the platform's own output. The paper's own thesis --- that machine-generated science
must carry receipts and honestly scoped verdicts --- was applied to the machine that generated it.

There is a replicability dividend here that we did not anticipate at the outset. The workbench exposes no
programmatic interface, so an analysis run in it is ordinarily a one-off, hand-paced web session. By driving
it headlessly and pulling its own provenance, we turned that one-off into a re-runnable, audited pipeline ---
and the sensitivity analyses in \S{}4.1b make the point concrete: the same panel, run locally and then inside
the workbench, reproduced byte-for-byte, verified by the workbench's own reviewer. An agentic analysis that
would otherwise be irreproducible-by-construction became replicable in principle --- subject to the
platform's user interface remaining stable (\S{}5.3). This is a small step, but a real one, toward
replicable-in-principle \emph{agentic} science rather than merely reproducible code.

\subsection*{5.2 The next experiment the approach nominates}\label{the-next-experiment-the-approach-nominates}

The value of a prioritization loop is not only that it culls, but that its survivors come with a sharp,
falsifiable next question. For NAB2, that question is directional. The Perturb-seq substrate tells us that
knocking NAB2 \emph{down} reshapes the Th1/Th2 program; it does not tell us whether, in disease, one would want
more NAB2 or less. We pursued that question orthogonally --- outside the Perturb-seq substrate --- and report the
result as a nomination for the next experiment, not as a finding.

Two lines of evidence bear on it. First, NAB2 is not a cancer dependency in the DepMap screens, which is
consistent with (though not proof of) a regulatory rather than an essential role --- non-contradictory context
rather than support. Second, and more suggestive, an \emph{exploratory} mining of public lesional-versus-non-lesional
atopic-dermatitis expression data (bulk and single-cell; datasets and accessions in the project's
direction-analysis record, and outside the referee substrate) for the \emph{direction} of the
NAB2--atopic-eczema association: NAB2 reads \textbf{down} in lesional atopic-dermatitis skin relative to non-lesional
(bulk log\textsubscript{2}FC -0.32, FDR 0.002; anti-correlated with Th2 activity, $\rho$ -0.34), and a single-cell analysis
confirmed this is a per-cell reduction in the disease-relevant compartments (keratinocyte -0.51,
paired \emph{p} 0.027; T/NK -0.57) rather than a shift in cell-type composition. Read together, these are more
consistent with NAB2 acting as a \textbf{Th2 brake that is lost or suppressed in chronic lesions} than as a Th2
driver --- a reading whose concordance with the perturbation data turns on the \emph{sign} of the Th2-program
change under NAB2 knockdown, which the substrate's directional metrics do not by themselves settle here, so
we advance the brake model as the directional \emph{question} to test rather than an established concordance. If that reading holds, the naive
therapeutic move --- topical knockdown --- is likely \emph{backwards}, and the direction to test is
\textbf{restoration or up-modulation} of NAB2.

We are candid about the ceiling on this. Expression tracks the direction of \emph{association}, not the
distinction between an effector and a brake; the voting-arm evidence was mixed (one significant arm, others
ambiguous or null); and only a perturbation --- showing that \emph{restoring} NAB2 dampens the Th2 program --- can
convert this brake hypothesis into a directional therapeutic claim. That perturbation is the experiment the
loop nominates. This is exactly the shape we argue for: the approach does not end the inquiry, it points a
bench at the one experiment most worth running.

\subsection*{5.3 Limitations}\label{limitations}

We inherit the LBD tradition of a candid limitations section, and there is a great deal to be candid about.

\noindent \textbf{One flagship, worked to depth; the rest nominated.} The full falsification apparatus --- the STAT6
\emph{cis}-test, the independent cross-model replication, and the three-way 12q13 decomposition --- was
applied to a single survivor (NAB2). The other 29 clean survivors carry their per-hop receipts but were not
individually adjudicated to that depth, so the confounder-falsification workflow is \emph{demonstrated}, not
\emph{evaluated at scale}; systematically applying it across survivors (and the held-out and external
benchmarks of \S5.3b) is the work that would turn demonstration into measured performance.

\noindent \textbf{The verdicts are nominations, not causal claims.} Every ``supported'' verdict means the chain held with a
receipt at each hop, not that the biology is established. This is sharpest at the disease hop, which is an
\emph{association} receipt: the module--disease labels derive from genetic (GWAS-based) evidence with no
colocalization or linkage-disequilibrium control, so NAB2's atopic-eczema label could in principle be
LD-inherited from the adjacent 12q13 atopy locus. We falsified the \emph{cis}-effect on STAT6's expression, which
addresses the perturbation signal; it does not settle the provenance of the GWAS label, which we foreground
as the flagship nomination's key open question rather than treat as discharged (\S{}4.4b), and which only
variant-level colocalization against the 12q13 atopic-dermatitis GWAS can resolve.

\noindent \textbf{The disease hop's stringency is substrate-inherited, not the referee's own discrimination.} Our
label-shuffle control (\S{}4.1b) is the most important honesty check in the paper, and it returned a result we
report straight: the disease hop supports fewer than 1\% of arbitrary gene$\times$disease pairs, but so does a
random relabeling of the enrichment data --- the near-total refutation rate is a property of the enrichment
study's FDR family, inherited by the referee, not a discrimination the referee itself supplies. The
referee's confident-\emph{no} edge lies in the knockdown-QC gate and the effect/program demotions, which
are its own computations; the disease hop's contribution is a stringent, label-dependent nomination filter,
and we claim no more for it.

\noindent \textbf{The funnel numbers are conditioned, not free-standing.} Within the pre-gated universe the program hop is
a tautology (\texttt{refuted\_program\ $\equiv$\ 0}), so it discriminates in an individual receipt, not as a funnel filter;
and ``43 supported'' is a joint product of the literature-novelty gate and the referee (the referee alone
supports 395 of 47,220 pairs). We restate these in Results rather than burying them here.

\noindent \textbf{The substrate is bounded, and singular.} The prioritization is retrospective against one Perturb-seq
resource, one program (Th1/Th2), one cell type, and one set of aggregated tables; we ran no new wet-lab
experiment. The approach is only as good as the held substrate for a given claim, and it stays silent ---
returning \emph{untested} --- where the substrate cannot speak, which is a feature, but also a boundary on scope.

\noindent \textbf{The determinism boundary, and where model judgment enters.} Every data receipt is deterministic code; the
language model interprets receipts and never computes one. But the objective's weights are a human design
choice (we showed the headline is robust to them, \S{}4.1b), and the model's interpretive and provenance-writing
roles are judgment, not computation --- so the calibrated-language discipline, and the independent critic that
enforces it, are load-bearing rather than cosmetic.

\noindent \textbf{The self-audit is within-family, and the automation depends on an unstable surface.} The workbench's
reviewer is an independent \emph{role, model, and checkpoint}, but within a single model family; cross-family
independence is supplied externally (\S{}4.6), not natively. And the headless driver depends on the workbench's
web front-end, for which there is no stable public contract --- a replicability bridge over an API-less
platform, honestly, not a durable interface. If the front-end changes, the driver must be re-fitted.

\subsection*{5.3b What would strengthen the evidence, and how}\label{strengthening-the-evidence}

We are direct about what the diagnostics of Section 4 do and do not establish. They show that the
\emph{machinery} behaves as designed: the quality-control gate converts every failed knockdown into
\emph{untested} without leakage, the referee refutes and demotes with receipts, its disease-hop stringency is
label-dependent, and the flagship's rank is stable to the analysis's free parameters. They do \emph{not}
establish that the referee \emph{decides correctly at scale} --- there is no external validation and no
precision or recall against a ground truth, and we make no such claim. That gap is not a flaw to bury but the
next programme of work, and the loop nominates it concretely.

Four steps, in priority order, would convert these behavioural diagnostics into a measured decision-quality
claim. \textbf{(1) A time-sliced held-out evaluation} \cite{yetisgen2009}: freeze the literature at a cutoff,
generate against the pre-cutoff corpus, and measure recovery of gene$\rightarrow$program$\rightarrow$disease
relationships established only afterward, reporting precision@$k$, mean average precision, and recall --- the
standard that separates ``the machinery runs'' from ``the machinery is right.'' \textbf{(2) An external
ground-truth panel} of known-true and known-false T-cell-program$\rightarrow$disease relationships, drawn from
curated association evidence \cite{mountjoy2021} and the immunology literature, scored for precision, recall,
and --- because the confident \emph{no} is the thesis --- the false-discovery and false-abstention rates on
the \emph{refuted} and \emph{untested} classes specifically. \textbf{(3) Colocalization for the flagship}: the
one confounder the substrate cannot discharge (\S4.4b) is exactly the analysis the loop nominates --- a
variant-level colocalization or summary-data Mendelian randomization test \cite{giambartolomei2014,smr2016}
between the 12q13 atopic-dermatitis GWAS \cite{paternoster2015} and CD4$^+$ T-cell \emph{NAB2} versus
\emph{STAT6} expression, decisive given the two genes' convergent geometry, and a direct instance of the
association-to-cause ladder that proximity alone cannot climb \cite{smemo2014,schaid2018}. \textbf{(4)
Prospective perturbation}, the field's emerging gold standard \cite{gottweis2025}: an independent NAB2
knockdown --- or, per the brake reading (\S5.2), a restoration --- in primary CD4$^+$ T cells, read out on
Th1/Th2 markers with an orthogonal method to control for CRISPRi \emph{cis}-artifacts. None of these is gating
for the present contribution --- a prioritization method with explicit abstention and falsification
diagnostics --- but each is a rung on the ladder from association to cause, and naming them precisely is part
of the calibrated-language discipline the paper argues for.

\subsection*{5.4 Conclusion}\label{conclusion}

Literature-based discovery has never lacked for hypotheses; it has lacked a disciplined way to decide which
ones to pursue, and the candor to say \emph{no}. Wayfinder pairs a machine hypothesis-generator with a
deterministic referee that evaluates each proposal against a held experimental substrate and returns a
prioritization verdict with a receipt at every hop --- supporting, refuting, or declining to test --- and runs the whole loop,
including an audit of its own language, inside an agentic workbench whose work we can reproduce and inspect.
Applied to a genome-scale CD4\textsuperscript{+} T-cell resource, it culled tens of thousands of machine-generated hypotheses
to a small set of receipt-backed survivors, of which NAB2 $\rightarrow$ Th1/Th2 $\rightarrow$ atopic eczema is the worked example: a
literature-novel regulatory nomination whose sharpest artifactual confounder we falsified against the study
authors' own data, and which points to a concrete, falsifiable next experiment. The contribution is the
method and what it found --- receipt-backed prioritization of machine-generated hypotheses, with explicit
abstention and falsification diagnostics --- and the confident, receipt-backed \emph{no} that a triage step
must be willing to give.
